Predicting Fire Risk: \\ An Evaluation of the Fire Danger Index

### 2016-01-22 paper.tex

McArthur's FFDI originated as tabulated indices and were subsequently reduced to equations by \citet{noble80} which in turn simplified into the formulation of the FFDI given by equation (\ref{eq FFDI}).

\begin{equation}
\label{eq FFDI}
F=2.0 e^{-0.450 + 0.987 \ln(D)-0.0345H+0.0338T+0.0234V}
\end{equation}
where 
\begin{itemize}
	\item $D$: drought factor (Index)
	\item $H$: relative humidity (percent)
	\item $T$: temperature (degrees Celsius)
	\item $V$: Average wind velocity (\emph{km/hr})
\end{itemize}

@@

\begin{table}[ht]
	\centering
	\begin{tabular}{|c|c|}
		\hline Fire Danger Rating & Forest Fire Danger Index \\ 
		\hline
		Low-Moderate  & 0-11 \\ 
		High  &  12-24 \\ 
		Very High  &  25-49\\ 
		Severe &  50-74\\ 
		Extreme  &  75-99 \\ 
		Code Red  &  100+ \\ 
		\hline 
	\end{tabular}
	\caption{The Forest Fire Danger Index and the Fire Danger Rating}
	\label{table:FDR}
\end{table}

@@

\begin{figure}[h]
	\centering 
%	\includegraphics[width=1\columnwidth]{figures/Data/cfadis.pdf} 
	\caption{Country Fire Authority Districts} 
	\label{fig:dist} 
\end{figure}

@@

%Graphs, need to redo better
\begin{figure}[h]
	\centering 
%	\includegraphics[width=1\columnwidth]{figures/Data/silvan.pdf} 
	\caption{Time series of meteorological variables for Silvan (-37.82, 145.42)} 
	\label{fig:silv} 
\end{figure}

@@

%WS, Temp and FDR contours, see if can change colours to match with the diagram
\begin{figure}[h]
	\centering 
%	\includegraphics[width=1\columnwidth]{figures/Data/tw_fdr.pdf} 
	\caption{Wind Speed and temperature coloured by fire danger rating} 
	\label{fig:tw_fdr} 
\end{figure}

@@

% level plot of GS curing, probably needs resizing
\begin{figure}[h]
	\centering 
%	\includegraphics[width=1\columnwidth]{figures/Data/cur_lp.pdf} 
	\caption{Level Plot of Grassland Curing } 
	\label{fig:cur_lp} 
\end{figure}

@@

% level plot of vapour pressure, level plot may be unnesc could just have std plot
\begin{figure}[h]
	\centering 
%	\includegraphics[width=1\columnwidth]{figures/Data/vap_ras.pdf} 
	\caption{Raster plot of vapour pressure 9am } 
	\label{fig:vap_ras} 
\end{figure}

@@

\begin{table}[ht]
	\centering
	\begin{tabular}{rrrrrrrrr}
		\hline
		& ffdi & temp & df & rh & ws & curing & rain & vap \\ 
		\hline
		ffdi & 1.00 & 0.77 & 0.62 & -0.77 & 0.04 & 0.51 & -0.20 & 0.01 \\ 
		temp & 0.77 & 1.00 & 0.43 & -0.70 & -0.18 & 0.46 & -0.10 & 0.39 \\ 
		df & 0.62 & 0.43 & 1.00 & -0.58 & -0.09 & 0.47 & -0.37 & -0.08 \\ 
		rh & -0.77 & -0.70 & -0.58 & 1.00 & 0.19 & -0.49 & 0.31 & 0.06 \\ 
		ws & 0.04 & -0.18 & -0.09 & 0.19 & 1.00 & 0.01 & 0.07 & -0.16 \\ 
		curing & 0.51 & 0.46 & 0.47 & -0.49 & 0.01 & 1.00 & -0.11 & 0.05 \\ 
		rain & -0.20 & -0.10 & -0.37 & 0.31 & 0.07 & -0.11 & 1.00 & 0.17 \\ 
		vap & 0.01 & 0.39 & -0.08 & 0.06 & -0.16 & 0.05 & 0.17 & 1.00 \\ 
		\hline
	\end{tabular}
	\caption{Correlation matrix of variables}
	\label{table:correl}
\end{table}

@@

%Commented out until final printing, 
%Map of fire ignitions in Victoria
\begin{figure}[h]
	\centering 
%	\includegraphics[width=1\columnwidth]{figures/data/firemap.pdf} 
	\caption{Map of fire ignition 2000 - 2009} 
	\label{fig:firemap} 
\end{figure}

%timeseries histogram with count of ignitions per fortnight
\begin{figure}[h]
	\centering 
%	\includegraphics[width=1\columnwidth]{figures/data/occurence.pdf} 
	\caption{Number of ignitions by fortnight} 
	\label{fig:occ} 
\end{figure}

@@

\begin{figure}[h]
	\centering 
%	\includegraphics[width=1\columnwidth]{figures/Data/tw_fire.pdf} 
	\caption{Wind Speed and temperature coloured by fire/no fire} 
	\label{fig:tw_fire} 
\end{figure}

@@

\begin{figure}[h]
	\centering 
%	\includegraphics[width=1\columnwidth]{figures/Data/dh_fire.pdf} 
	\caption{Drought factor and relative humidity coloured by fire/no fire} 
	\label{fig:dh_fire} 
\end{figure}

@@

\begin{figure}[h]
	\centering 
%	\includegraphics[width=1\columnwidth]{figures/Data/gv_fire.pdf} 
	\caption{Grassland curing and vapour pressure coloured by fire/no fire} 
	\label{fig:gv_fire} 
\end{figure}

@@

\begin{enumerate}
	\item Take a bootstrap sample (random sample of size N with replacement) from the data. 
	\item Take a random sample of size $M \textless P$ (typically $M \approx \sqrt{P}$, \cite{james13}, without replacement) from the predictors. 
	\item Construct a classification tree using these samples of data and predictors.
	\item Repeat from step 2 for each subsequent split until the tree is as large as desired. Do not prune. 
	\item Drop observations not included in the bootstrap sample from step 1 (known as out-of-bag data) down the tree. The class assigned to each observation is stored along with each observations predictor values.
	\item Repeat steps 1-5 a large number of times to grow a forest of trees. 
	\item In order to determine the classification of a new observation. Have each tree make a classification and use a majority vote for the final prediction.
\end{enumerate}

@@

For consistency all three of the classification techniques are estimated on the balanced training set and evaluated by  on the test data. I evaluate based on receiver operating characteristic (ROC) plots  and using 2 x 2 confusion matrices.

An example confusion matrix is given in table \ref{table:ex}. 
%eg confusion matrix with ABCD
\begin{table}[ht]
	\centering
	\begin{tabular}{r|rr|r}
		& Model & & \\ 
		\hline
		& Prediction   & &  \\ 
		Reference & No Fire & Fire & \\ 
		\hline
		No Fire & a & a & a+b \\ 
		Fire & c & d  & c + d \\ 
		\hline
		 & a+c & b+d & N \\
	\end{tabular}
	\caption{Example confusion matrix}
	\label{table:ex}
\end{table}

@@

I also report sensitivity (correctly classified fires divided by total number of actual fires) and specificity (correctly identified non-fires divided by total number of non-fires), balanced accuracy and kappa.

\begin{equation}
\label{eq sens}
\mbox{Sensitivity} = \frac{a }{a+b}
\end{equation}

\begin{equation}
\label{eq spec}
\mbox{Specificity} = \frac{d}{c+d}
\end{equation}

\begin{equation}
\label{eq BA}
\mbox{Balanced Accuracy} = \frac{\frac{a }{a+b} + \frac{d}{c+d} }{2}
\end{equation}

\begin{equation}
\label{eq:K}
\kappa = \frac{P_0 - P_c}{1-P_c}
\end{equation}

where $P_0$ is the total agreement probability and $P_c$ is the Agreement probability which is due to chance (see \cite{kappa} for definitions).

On top of the accuracy measures in equations (\ref{eq sens}), (\ref{eq spec}) and (\ref{eq BA}) the value in the bottom left of the confusion matrix (c) where we incorrectly predict ~No fire' is very important in this context.

@@

%Output from best performing splines model (used xtable rather than texreg, too inconsistent?)
\begin{table}[ht]
	\centering
	\begin{tabular}{rrrrr}
		\hline
		& Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
		\hline
		(Intercept) & -3.1484 & 0.2757 & -11.4210 & 0.0000 \\ 
		ns(ffdi, 5)1 & -0.2905 & 0.3613 & -0.8043 & 0.4213 \\ 
		ns(ffdi, 5)2 & -0.2304 & 0.4625 & -0.4982 & 0.6183 \\ 
		ns(ffdi, 5)3 & 0.4040 & 0.5589 & 0.7229 & 0.4697 \\ 
		ns(ffdi, 5)4 & -0.4340 & 0.9674 & -0.4487 & 0.6537 \\ 
		ns(ffdi, 5)5 & -0.8815 & 0.7271 & -1.2124 & 0.2254 \\ 
		ns(temp, 1) & 1.8543 & 0.3416 & 5.4276 & 0.0000 \\ 
		ns(rh, 2)1 & 0.9325 & 0.5687 & 1.6397 & 0.1011 \\ 
		ns(rh, 2)2 & -0.6173 & 0.3607 & -1.7115 & 0.0870 \\ 
		ws & -0.0169 & 0.0146 & -1.1541 & 0.2485 \\ 
		ns(df, 5)1 & 1.1093 & 0.2798 & 3.9639 & 0.0001 \\ 
		ns(df, 5)2 & 1.0156 & 0.3115 & 3.2601 & 0.0011 \\ 
		ns(df, 5)3 & 0.7113 & 0.2604 & 2.7319 & 0.0063 \\ 
		ns(df, 5)4 & 0.8071 & 0.5479 & 1.4730 & 0.1408 \\ 
		ns(df, 5)5 & -0.1512 & 0.1950 & -0.7752 & 0.4382 \\ 
		ns(curing, 6)1 & 1.5648 & 0.1388 & 11.2754 & 0.0000 \\ 
		ns(curing, 6)2 & 2.3986 & 0.1874 & 12.7997 & 0.0000 \\ 
		ns(curing, 6)3 & 1.6637 & 0.1640 & 10.1436 & 0.0000 \\ 
		ns(curing, 6)4 & 1.6949 & 0.1335 & 12.6921 & 0.0000 \\ 
		ns(curing, 6)5 & 1.4634 & 0.3436 & 4.2588 & 0.0000 \\ 
		ns(curing, 6)6 & -0.4909 & 0.0817 & -6.0066 & 0.0000 \\ 
		rain & -0.0286 & 0.0056 & -5.0703 & 0.0000 \\ 
		vap & -0.0003 & 0.0067 & -0.0504 & 0.9598 \\ 
		\hline
	\end{tabular}
	\caption{Regression output for Model 4}
	\label{table:logf}
\end{table}

%GAM plots, not sure if this is exactly splines... 
\begin{figure}[h]
\null\vfill\noindent
%\includegraphics[page=1,width=.48\linewidth,height=.45\textheight,keepaspectratio]{figures/gam_pl.pdf}
\hfill
%\includegraphics[page=2,width=.48\linewidth,height=.45\textheight,keepaspectratio]{figures/gam_pl.pdf}
\vfill
%\includegraphics[page=3,width=.48\linewidth,height=.45\textheight,keepaspectratio]{figures/gam_pl.pdf}
\hfill
%\includegraphics[page=4,width=.48\linewidth,height=.45\textheight,keepaspectratio]{figures/gam_pl.pdf}
\vfill
%\includegraphics[page=5,width=.48\linewidth,height=.45\textheight,keepaspectratio]{figures/gam_pl.pdf}
\hfill
%\includegraphics[page=6,width=.48\linewidth,height=.45\textheight,keepaspectratio]{figures/gam_pl.pdf}
\vfill
\null
	\caption{Plots of smoothing functions in Model 4 with 95\% point-wise confident interval} 
	\label{fig:gam_pl} 
\end{figure}

@@

%Confusion matrix cutoff=0.5
\begin{table}[ht]
	\centering
	\begin{tabular}{r|rr|rr|rr|rr}
		\hline
		& Model 1 &  & Model 2 &  & Model 3 &  & Model 4 & \\ 
		\hline
		& Actual   &  &  &  &  & & & \\ 
		Predict & No Fire & Fire & No Fire & Fire & No Fire & Fire & No Fire & Fire\\ 
		\hline
No Fire & 0.5959 & 0.0030 & 0.4457 & 0.0019 & 0.5111 & 0.0021 & 0.5807 & 0.0016 \\ 
Fire & 0.3989 & 0.0022 & 0.5490 & 0.0034 & 0.4838 & 0.0030 & 0.4142 & 0.0035 \\ 
		\hline
	\end{tabular}
	\caption{Confusion Matrices for Models 1, 2 \& 3 cutoff value 0.5}
	\label{table:cm2}
\end{table}

% Accuracy for cutoff 0.5
\begin{table}[ht]
	\centering
	\begin{tabular}{rrrrr}
		\hline
		& 1 & 2 & 3 & 4 \\ 
		\hline
Sensitivity & 0.5990 & 0.4481 & 0.5137 & 0.5837 \\ 
Specificity & 0.4267 & 0.6476 & 0.5854 & 0.6920 \\ 
Balanced Accuracy & 0.5129 & 0.5479 & 0.5495 & 0.6378 \\ 
Accuracy & 0.5981 & 0.4492 & 0.5141 & 0.5842 \\ 
Kappa & 0.0007 & 0.0018 & 0.0021 & 0.0067 \\
		\hline
	\end{tabular}
		\caption{Accuracy for Models 1, 2, 3 \& 4 cutoff value 0.5}
		\label{table:acc2}
\end{table}

@@

%ROC for logits
\begin{figure}[h]
	\centering 
%	\includegraphics[width=1\columnwidth]{figures/all_roc.pdf} 
	\caption{ROC curve for logistic models} 
	\label{fig:ffdi} 
\end{figure}

@@

% Trees with all variables
\begin{figure}[h]
	\centering 
%	\includegraphics[page=1, width=1\columnwidth]{figures/trees2.pdf} 
%	\includegraphics[page=2, width=1\columnwidth]{figures/trees2.pdf} 
	\caption{Trees with all variables} 
	\label{fig:tree2} 
\end{figure}

@@

%Confusion matrix cutoff=0.5
\begin{table}[ht]
	\centering
	\begin{tabular}{r|rr|rr}
		\hline
		  & Tree 1 &  & Tree 2 & \\ 
		\hline
		& Actual   &  &  &   \\ 
		Predict  & No Fire & Fire & No Fire & Fire\\ 
		\hline
No Fire  & 0.5553 & 0.0017 & 0.5088 & 0.0014 \\ 
Fire  & 0.4394 & 0.0036 & 0.4859 & 0.0038 \\ 
		\hline
	\end{tabular}
	\caption{Confusion Matrices for trees cutoff value 0.5}
	\label{table:tcm2}
\end{table}

% Accuracy for cutoff 0.5
\begin{table}[H]
	\centering
	\begin{tabular}{rrr}
		\hline
		 & Tree 1 & Tree 2 \\ 
		\hline
		Sensitivity  & 0.5582 & 0.5115 \\ 
		Specificity & 0.6832 & 0.7273 \\  
		Balanced Accuracy  & 0.6207 & 0.6194 \\ 
		Accuracy  & 0.5589 & 0.5127 \\ 
		Kappa & 0.0055 & 0.0051 \\ 
		\hline
	\end{tabular}
	\caption{Accuracy for trees cutoff value 0.5}
	\label{table:tacc2}
\end{table}

@@

\begin{figure}[h]
	\centering 
%	\includegraphics[width=1\columnwidth]{figures/trees_roc.pdf} 
	\caption{ROC curve for Tree Models} 
	\label{fig:troc} 
\end{figure}

@@

%confusion matrix on 0.5 cutoff value
\begin{table}[ht]
	\centering
	\begin{tabular}{rrr}
		\hline
		& Random Forest &  \\ 
		\hline
		& Actual   &   \\ 
		Predict  & No Fire & Fire \\ 
		\hline
		No Fire & 0.6093 & 0.0016 \\ 
		Fire & 0.3856 & 0.0035 \\ 
		\hline
	\end{tabular}
		\caption{Confusion matrix for random forest cutoff value 0.5}
		\label{table:cmrf}
\end{table}

\begin{table}[ht]
	\centering
	\begin{tabular}{rr}
		\hline
		& Random Forest \\ 
		\hline
		Sensitivity & 0.6124 \\ 
		Specificity & 0.6771 \\ 
		Balanced Accuracy & 0.6447 \\ 
		Accuracy & 0.6127 \\ 
		Kappa & 0.0075 \\ 
		\hline
	\end{tabular}
			\caption{Accuracy measure random forest cutoff value 0.5}
			\label{table:accrf}
\end{table}

@@

\begin{figure}[h]
	\centering 
%	\includegraphics[width=1\columnwidth]{figures/varimp.pdf} 
	\caption{Variable Importance for random forest} 
	\label{fig:vimp} 
\end{figure}

@@

\begin{table}[h]
	\centering
	\begin{tabular}{r|rrrr}
		\hline
		& Model 3 & Model 4 & Tree & Forest \\ 
		\hline
		AUC & 0.58 & 0.69 & 0.65 & 0.70 \\ 
		\hline
	\end{tabular}
	\caption{Areas under ROC curve for logistic model, classification trees \& forest}
	\label{table:AUC}
\end{table}

\begin{figure}[h]
	\centering 
%	\includegraphics[width=1\columnwidth]{figures/roc_rf.pdf} 
	\caption{Variable Importance for random forest} 
	\label{fig:rocrf} 
\end{figure}

>>>>

McArthur's FFDI originated as tabulated indices and were subsequently reduced to equations by \citet{noble80} which in turn simplified into the formulation of the FFDI given by
\begin{equation}
\label{eq FFDI}
F=2.0 e^{-0.450 + 0.987 \ln(D)-0.0345H+0.0338T+0.0234V},
\end{equation}
where 
\begin{itemize}
	\item $D$ is the drought factor index;
	\item $H$ is relative humidity (as a percentage);
	\item $T$ is temperature (degrees Celsius); and
	\item $V$ is average wind velocity (in km/hr).
\end{itemize}

@@

\begin{table}[!ht]
	\centering
	\begin{tabular}{ll}
		\toprule
		Fire Danger Rating & Forest Fire Danger Index \\
		\midrule
		Low-Moderate              & 0--11 \\
		High                      & 12--24 \\
		Very High                 & 25--49\\
		Severe                    & 50--74\\
		Extreme                   & 75--99 \\
		Code Red                  & 100+ \\
		\bottomrule
	\end{tabular}
	\caption{The Forest Fire Danger Index and the Fire Danger Rating}
	\label{table:FDR}
\end{table}

@@

\begin{figure}[!ht]
	\centering 
%	\includegraphics[width=1\columnwidth]{figures/Data/cfadis.pdf} 
	\caption{Country Fire Authority Districts} 
	\label{fig:dist} 
\end{figure}

@@

%Graphs, need to redo better
\begin{figure}[!ht]
	\centering 
%	\includegraphics[width=1\columnwidth]{figures/Data/silvan.pdf} 
	\caption{Time series of meteorological variables for Silvan (-37.82, 145.42)} 
	\label{fig:silv} 
\end{figure}

@@

%WS, Temp and FDR contours, see if can change colours to match with the diagram
\begin{figure}[!ht]
	\centering 
%	\includegraphics[width=1\columnwidth]{figures/Data/tw_fdr.pdf} 
	\caption{Wind Speed and temperature coloured by fire danger rating} 
	\label{fig:tw_fdr} 
\end{figure}

@@

% level plot of GS curing, probably needs resizing
\begin{figure}[!ht]
	\centering 
%	\includegraphics[width=1\columnwidth]{figures/Data/cur_lp.pdf} 
	\caption{Level Plot of Grassland Curing } 
	\label{fig:cur_lp} 
\end{figure}

@@

% level plot of vapour pressure, level plot may be unnesc could just have std plot
\begin{figure}[!ht]
	\centering 
%	\includegraphics[width=1\columnwidth]{figures/Data/vap_ras.pdf} 
	\caption{Raster plot of vapour pressure 9am } 
	\label{fig:vap_ras} 
\end{figure}

@@

\begin{table}[!ht]
	\centering
	\begin{tabular}{lrrrrrrrr}
		\toprule
		       & ffdi  & temp  & df    & rh    & ws    & curing & rain  & vap \\
		\midrule
		ffdi   & 1.00  & 0.77  & 0.62  & -0.77 & 0.04  & 0.51   & -0.20 & 0.01 \\
		temp   & 0.77  & 1.00  & 0.43  & -0.70 & -0.18 & 0.46   & -0.10 & 0.39 \\
		df     & 0.62  & 0.43  & 1.00  & -0.58 & -0.09 & 0.47   & -0.37 & -0.08 \\
		rh     & -0.77 & -0.70 & -0.58 & 1.00  & 0.19  & -0.49  & 0.31  & 0.06 \\
		ws     & 0.04  & -0.18 & -0.09 & 0.19  & 1.00  & 0.01   & 0.07  & -0.16 \\
		curing & 0.51  & 0.46  & 0.47  & -0.49 & 0.01  & 1.00   & -0.11 & 0.05 \\
		rain   & -0.20 & -0.10 & -0.37 & 0.31  & 0.07  & -0.11  & 1.00  & 0.17 \\
		vap    & 0.01  & 0.39  & -0.08 & 0.06  & -0.16 & 0.05   & 0.17  & 1.00 \\
		\bottomrule
	\end{tabular}
	\caption{Correlation matrix of variables}
	\label{table:correl}
\end{table}

@@

%Commented out until final printing, 
%Map of fire ignitions in Victoria
\begin{figure}[!ht]
	\centering 
%	\includegraphics[width=1\columnwidth]{figures/data/firemap.pdf} 
	\caption{Map of fire ignition 2000 - 2009} 
	\label{fig:firemap} 
\end{figure}

%timeseries histogram with count of ignitions per fortnight
\begin{figure}[!ht]
	\centering 
%	\includegraphics[width=1\columnwidth]{figures/data/occurence.pdf} 
	\caption{Number of ignitions by fortnight} 
	\label{fig:occ} 
\end{figure}

@@

\begin{figure}[!ht]
	\centering 
%	\includegraphics[width=1\columnwidth]{figures/Data/tw_fire.pdf} 
	\caption{Wind Speed and temperature coloured by fire/no fire} 
	\label{fig:tw_fire} 
\end{figure}

@@

\begin{figure}[!ht]
	\centering 
%	\includegraphics[width=1\columnwidth]{figures/Data/dh_fire.pdf} 
	\caption{Drought factor and relative humidity coloured by fire/no fire} 
	\label{fig:dh_fire} 
\end{figure}

@@

\begin{figure}[!ht]
	\centering 
%	\includegraphics[width=1\columnwidth]{figures/Data/gv_fire.pdf} 
	\caption{Grassland curing and vapour pressure coloured by fire/no fire} 
	\label{fig:gv_fire} 
\end{figure}

@@

\begin{enumerate}
	\item Take a bootstrap sample (random sample of size N with replacement) from the data. 
	\item Take a random sample of size $M < P$ (typically $M \approx \sqrt{P}$, \cite{james13}, without replacement) from the predictors. 
	\item Construct a classification tree using these samples of data and predictors.
	\item Repeat from step 2 for each subsequent split until the tree is as large as desired. Do not prune. 
	\item Drop observations not included in the bootstrap sample from step 1 (known as out-of-bag data) down the tree. The class assigned to each observation is stored along with each observations predictor values.
	\item Repeat steps 1-5 a large number of times to grow a forest of trees. 
	\item In order to determine the classification of a new observation. Have each tree make a classification and use a majority vote for the final prediction.
\end{enumerate}

@@

For consistency all three of the classification techniques are estimated on the balanced training set and evaluated by  on the test data. I evaluate based on receiver operating characteristic (ROC) plots  and using $2\times2$ confusion matrices.

An example confusion matrix is given in Table~\ref{table:ex}. 
%eg confusion matrix with ABCD
\begin{table}[!ht]
	\centering
	\begin{tabular}{lr|rr|r|}
    & 	\multicolumn{1}{l}{} & \multicolumn{2}{c}{\emph{Model Prediction}} \\
    & 	\multicolumn{1}{l}{} & \multicolumn{1}{r}{No Fire} & \multicolumn{1}{r}{Fire} \\
		\cline{3-5}
\emph{Actual} & No Fire & a   & a   & a+b \\
                 & 		Fire & c   & d   & c + d \\
		\cline{3-5}
                 & 		     & a+c & b+d & N \\
		\cline{3-5}
	\end{tabular}
	\caption{Example confusion matrix}
	\label{table:ex}
\end{table}

@@

I also report sensitivity (correctly classified fires dividied by total number of actual fires) and specificity (correctly identified non-fires divided by total number of non-fires), balanced accuracy and kappa:
\begin{align}
\label{eq sens}
\mbox{Sensitivity} &= \frac{a }{a+b} \\
\label{eq spec}
\mbox{Specificity} &= \frac{d}{c+d} \\
\label{eq BA}
\mbox{Balanced Accuracy} &= \frac{\frac{a }{a+b} + \frac{d}{c+d} }{2} \\
\label{eq:K}
\kappa &= \frac{P_0 - P_c}{1-P_c}
\end{align}
where $P_0$ is the total agreement probability and $P_c$ is the Agreement probability which is due to chance (see \cite{kappa} for definitions).

On top of the accuracy measures in equations (\ref{eq sens}), (\ref{eq spec}) and (\ref{eq BA}) the value in the bottom left of the confusion matrix (c) where we incorrectly predict `No fire' is very important in this context.

@@

%Output from best performing splines model (used xtable rather than texreg, too inconsistent?)
\begin{table}[!ht]
	\centering
	\begin{tabular}{rrrrr}
		\toprule
		& Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
		\midrule
		(Intercept) & -3.1484 & 0.2757 & -11.4210 & 0.0000 \\ 
		ns(ffdi, 5)1 & -0.2905 & 0.3613 & -0.8043 & 0.4213 \\ 
		ns(ffdi, 5)2 & -0.2304 & 0.4625 & -0.4982 & 0.6183 \\ 
		ns(ffdi, 5)3 & 0.4040 & 0.5589 & 0.7229 & 0.4697 \\ 
		ns(ffdi, 5)4 & -0.4340 & 0.9674 & -0.4487 & 0.6537 \\ 
		ns(ffdi, 5)5 & -0.8815 & 0.7271 & -1.2124 & 0.2254 \\ 
		ns(temp, 1) & 1.8543 & 0.3416 & 5.4276 & 0.0000 \\ 
		ns(rh, 2)1 & 0.9325 & 0.5687 & 1.6397 & 0.1011 \\ 
		ns(rh, 2)2 & -0.6173 & 0.3607 & -1.7115 & 0.0870 \\ 
		ws & -0.0169 & 0.0146 & -1.1541 & 0.2485 \\ 
		ns(df, 5)1 & 1.1093 & 0.2798 & 3.9639 & 0.0001 \\ 
		ns(df, 5)2 & 1.0156 & 0.3115 & 3.2601 & 0.0011 \\ 
		ns(df, 5)3 & 0.7113 & 0.2604 & 2.7319 & 0.0063 \\ 
		ns(df, 5)4 & 0.8071 & 0.5479 & 1.4730 & 0.1408 \\ 
		ns(df, 5)5 & -0.1512 & 0.1950 & -0.7752 & 0.4382 \\ 
		ns(curing, 6)1 & 1.5648 & 0.1388 & 11.2754 & 0.0000 \\ 
		ns(curing, 6)2 & 2.3986 & 0.1874 & 12.7997 & 0.0000 \\ 
		ns(curing, 6)3 & 1.6637 & 0.1640 & 10.1436 & 0.0000 \\ 
		ns(curing, 6)4 & 1.6949 & 0.1335 & 12.6921 & 0.0000 \\ 
		ns(curing, 6)5 & 1.4634 & 0.3436 & 4.2588 & 0.0000 \\ 
		ns(curing, 6)6 & -0.4909 & 0.0817 & -6.0066 & 0.0000 \\ 
		rain & -0.0286 & 0.0056 & -5.0703 & 0.0000 \\ 
		vap & -0.0003 & 0.0067 & -0.0504 & 0.9598 \\ 
		\bottomrule
	\end{tabular}
	\caption{Regression output for Model 4}
	\label{table:logf}
\end{table}

%GAM plots, not sure if this is exactly splines... 
\begin{figure}[!ht]
\null\vfill\noindent
%\includegraphics[page=1,width=.48\linewidth,height=.45\textheight,keepaspectratio]{figures/gam_pl.pdf}
\hfill
%\includegraphics[page=2,width=.48\linewidth,height=.45\textheight,keepaspectratio]{figures/gam_pl.pdf}
\vfill
%\includegraphics[page=3,width=.48\linewidth,height=.45\textheight,keepaspectratio]{figures/gam_pl.pdf}
\hfill
%\includegraphics[page=4,width=.48\linewidth,height=.45\textheight,keepaspectratio]{figures/gam_pl.pdf}
\vfill
%\includegraphics[page=5,width=.48\linewidth,height=.45\textheight,keepaspectratio]{figures/gam_pl.pdf}
\hfill
%\includegraphics[page=6,width=.48\linewidth,height=.45\textheight,keepaspectratio]{figures/gam_pl.pdf}
\vfill
\null
	\caption{Plots of smoothing functions in Model 4 with 95\% point-wise confident interval} 
	\label{fig:gam_pl} 
\end{figure}

@@

%Confusion matrix cutoff=0.5
\begin{table}[!ht]
	\centering
	\begin{tabular}{r|rr|rr|rr|rr}
		\toprule
		& Model 1 &  & Model 2 &  & Model 3 &  & Model 4 & \\ 
		\midrule
		& Actual   &  &  &  &  & & & \\ 
		Predict & No Fire & Fire & No Fire & Fire & No Fire & Fire & No Fire & Fire\\ 
		\midrule
No Fire & 0.5959 & 0.0030 & 0.4457 & 0.0019 & 0.5111 & 0.0021 & 0.5807 & 0.0016 \\ 
Fire & 0.3989 & 0.0022 & 0.5490 & 0.0034 & 0.4838 & 0.0030 & 0.4142 & 0.0035 \\ 
		\bottomrule
	\end{tabular}
	\caption{Confusion Matrices for Models 1, 2 \& 3 cutoff value 0.5}
	\label{table:cm2}
\end{table}

% Accuracy for cutoff 0.5
\begin{table}[!ht]
	\centering
	\begin{tabular}{rrrrr}
		\toprule
		& 1 & 2 & 3 & 4 \\ 
		\midrule
Sensitivity & 0.5990 & 0.4481 & 0.5137 & 0.5837 \\ 
Specificity & 0.4267 & 0.6476 & 0.5854 & 0.6920 \\ 
Balanced Accuracy & 0.5129 & 0.5479 & 0.5495 & 0.6378 \\ 
Accuracy & 0.5981 & 0.4492 & 0.5141 & 0.5842 \\ 
Kappa & 0.0007 & 0.0018 & 0.0021 & 0.0067 \\
		\bottomrule
	\end{tabular}
		\caption{Accuracy for Models 1, 2, 3 \& 4 cutoff value 0.5}
		\label{table:acc2}
\end{table}

@@

%ROC for logits
\begin{figure}[!ht]
	\centering 
%	\includegraphics[width=1\columnwidth]{figures/all_roc.pdf} 
	\caption{ROC curve for logistic models} 
	\label{fig:ffdi} 
\end{figure}

@@

% Trees with all variables
\begin{figure}[!ht]
	\centering 
%	\includegraphics[page=1, width=1\columnwidth]{figures/trees2.pdf} 
%	\includegraphics[page=2, width=1\columnwidth]{figures/trees2.pdf} 
	\caption{Trees with all variables} 
	\label{fig:tree2} 
\end{figure}

@@

%Confusion matrix cutoff=0.5
\begin{table}[!ht]
	\centering
	\begin{tabular}{r|rr|rr}
		\toprule
		  & Tree 1 &  & Tree 2 & \\ 
		\midrule
		& Actual   &  &  &   \\ 
		Predict  & No Fire & Fire & No Fire & Fire\\ 
		\midrule
No Fire  & 0.5553 & 0.0017 & 0.5088 & 0.0014 \\ 
Fire  & 0.4394 & 0.0036 & 0.4859 & 0.0038 \\ 
		\bottomrule
	\end{tabular}
	\caption{Confusion Matrices for trees cutoff value 0.5}
	\label{table:tcm2}
\end{table}

% Accuracy for cutoff 0.5
\begin{table}[!ht]
	\centering
	\begin{tabular}{rrr}
		\toprule
		 & Tree 1 & Tree 2 \\ 
		\midrule
		Sensitivity  & 0.5582 & 0.5115 \\ 
		Specificity & 0.6832 & 0.7273 \\  
		Balanced Accuracy  & 0.6207 & 0.6194 \\ 
		Accuracy  & 0.5589 & 0.5127 \\ 
		Kappa & 0.0055 & 0.0051 \\ 
		\bottomrule
	\end{tabular}
	\caption{Accuracy for trees cutoff value 0.5}
	\label{table:tacc2}
\end{table}

@@

\begin{figure}[!ht]
	\centering 
%	\includegraphics[width=1\columnwidth]{figures/trees_roc.pdf} 
	\caption{ROC curve for Tree Models} 
	\label{fig:troc} 
\end{figure}

@@

%confusion matrix on 0.5 cutoff value
\begin{table}[!ht]
	\centering
	\begin{tabular}{rrr}
		\toprule
		& Random Forest &  \\ 
		\midrule
		& Actual   &   \\ 
		Predict  & No Fire & Fire \\ 
		\midrule
		No Fire & 0.6093 & 0.0016 \\ 
		Fire & 0.3856 & 0.0035 \\ 
		\bottomrule
	\end{tabular}
		\caption{Confusion matrix for random forest cutoff value 0.5}
		\label{table:cmrf}
\end{table}

\begin{table}[!ht]
	\centering
	\begin{tabular}{rr}
		\toprule
		& Random Forest \\ 
		\midrule
		Sensitivity & 0.6124 \\ 
		Specificity & 0.6771 \\ 
		Balanced Accuracy & 0.6447 \\ 
		Accuracy & 0.6127 \\ 
		Kappa & 0.0075 \\ 
		\bottomrule
	\end{tabular}
			\caption{Accuracy measure random forest cutoff value 0.5}
			\label{table:accrf}
\end{table}

@@

\begin{figure}[!ht]
	\centering 
%	\includegraphics[width=1\columnwidth]{figures/varimp.pdf} 
	\caption{Variable Importance for random forest} 
	\label{fig:vimp} 
\end{figure}

@@

\begin{table}[!ht]
	\centering
	\begin{tabular}{r|rrrr}
		\toprule
		& Model 3 & Model 4 & Tree & Forest \\ 
		\midrule
		AUC & 0.58 & 0.69 & 0.65 & 0.70 \\ 
		\midrule
	\end{tabular}
	\caption{Areas under ROC curve for logistic model, classification trees \& forest}
	\label{table:AUC}
\end{table}

\begin{figure}[!ht]
	\centering 
%	\includegraphics[width=1\columnwidth]{figures/roc_rf.pdf} 
	\caption{Variable Importance for random forest} 
	\label{fig:rocrf} 
\end{figure}
